The Impact of Alternation

### 2017-05-05 draft.tex
ok

Formally, let $\preds$ be a finite set of formulae with variables from
$Q \cup \vec{x}$, such that $\bot \in \preds$ and the variables in $Q$
can occur negated in some $\phi \in \preds$. We assume that $\preds$
is fixed for now, and refer to these formulae as \emph{predicates}.
Given a formula $\phi$, we denote by $\abs{\phi} \equiv \bigwedge
\set{\pi \in \preds \mid \phi \models \pi}$ the abstraction of $\phi$
w.r.t. the predicates in $\preds$. The abstract version of the
post-image and acceptance condition are defined as follows:
{\footnotesize\[ \abs{\Post{\A}}(\phi,\varepsilon) \equiv \phi
\hspace*{2mm}
\abs{\Post{\A}}(\phi,ua) \equiv \abs{(\Post{\A}(\abs{\Post{\A}}(\phi,u),a))} 
\hspace*{2mm}
\abs{\Accept{\A}}(u) \equiv \abs{\Post{\A}}(\iota,u) \wedge\!\!\!\bigwedge_{q \in Q \setminus F} 
(q \rightarrow \bot)\]}
\begin{lemma}\label{lemma:abstract-invariant}
  For any bijection $\mu : \nat \rightarrow \Sigma^*$, there exists
  $k>0$ such that $\bigvee_{i=1}^k \abs{\Post{\A}}(\iota,\mu(i))$ defines
  an invariant for $\A$. 
\end{lemma}
\proof{ It is sufficient to show that there exists $k\geq0$ such that for
  all $u \in \Sigma^*$ there exists $i \in [0,k]$ such that
  $\Post{\A}(\iota,u) \models \abs{\Post{\A}}(\iota,\mu(i))$. We have
  $\Post{\A}(\iota,u) \models \abs{\Post{\A}}(\iota,u)$ for all $u \in
  \Sigma^*$. But since $\preds$ is a finite set, also the set
  $\{\abs{\Post{\A}}(\iota,u) \mid u \in \Sigma^*\}$ is finite. Thus
  there exists $k \geq 0$ such that, for all $u \in \Sigma^*$ there
  exists $i \in [0,k]$ such that $\abs{\Post{\A}}(\iota,u) \eqiff
  \abs{\Post{\A}}(\iota,\mu(i))$, which concludes the proof. \qed}

We are left with fulfilling the second point from the definition of a
safety invariant. To this end, suppose that, for a given set $\preds$
of predicates, the invariant $\abs{\Inv}$, defined by the previous
lemma, meets point (\ref{it1:inv}) but not point (\ref{it2:inv}) from
the definition of invariants, where $\Post{\A}$ and $\Accept{\A}$
replace $\abs{\Post{\A}}$ and $\abs{\Accept{\A}}$, respectively. In
other words, there exists a finite sequence $u \in \Sigma^*$ such that
$\nu \in \abs{\Inv}(\beta)$ and $\beta\cup\nu \models
\abs{\Accept{\A}}(u)$, for some boolean $\beta : Q \rightarrow \booli$
and data $\nu : \vec{x} \rightarrow \Data^\I$ valuations. Such
a $u\in\Sigma^*$ is called a \emph{counterexample}.

Once a counterexample $u$ is discovered, there are two
possibilities. Either\begin{inparaenum}[(i)]
\item $\Accept{\A}(u)$ is satisfiable, in which case $L(\A) \neq \emptyset$, or
\item $\Accept{\A}(u)$ is unsatisfiable, in which case $u$ is called
  \emph{spurious}.
\end{inparaenum}
In the first case, our semi-algorithm stops and returns a witness for
non-emptiness, obtained from the satisfying valuation of
$\Accept{\A}(u)$ and in the second case, we must strenghten the
invariant by excluding from $\abs{\Inv}$ all pairs $(\beta,\nu)$ such
that $\beta\cup\nu \models \abs{\Accept{\A}}(u)$. This strenghtening
is carried out by adding to $\preds$ several predicates that are
sufficient to exclude the spurious counterexample.

@@

\begin{lemma}\label{lemma:progress}
  Given a non-empty sequence of input events $u \in \Sigma^*$ such
  that $\Accept{\A}(u)$ is unsatisfiable, then the following
  hold: \begin{compactenum}
  \item\label{it1:progress} $\pi(u)$ is unsatisfiable, and
    %
  \item\label{it2:progress} if $\tuple{\top,I_1,\dots,I_n,\bot}$ is an
    interpolant for $\pi(u)$ and $I_i \in \preds$, for all
    $i\in[1,n]$, then $\abs{\Accept{\A}}(u)$ is unsatisfiable.
  \end{compactenum}
\end{lemma}
\proof{

>>>>

Formally, let $\preds$ be a finite set of formulae with variables from
$Q \cup \vec{x}$, such that $\bot \in \preds$ and the variables in $Q$
can occur negated in some $\phi \in \preds$. We assume that $\preds$
is fixed for now, and refer to these formulae as \emph{predicates}.
Given a formula $\phi$, we denote by $\abs{\phi} \equiv \bigwedge
\set{\pi \in \preds \mid \phi \models \pi}$ the abstraction of $\phi$
w.r.t. the predicates in $\preds$. The abstract version of the
post-image and acceptance condition are defined as follows:
{\footnotesize\[ \AbsPost{\A}(\phi,\varepsilon) \equiv \phi
\hspace*{2mm}
\AbsPost{\A}(\phi,ua) \equiv \abs{(\Post{\A}(\AbsPost{\A}(\phi,u),a))} 
\hspace*{2mm}
\AbsAccept{\A}(u) \equiv \AbsPost{\A}(\iota,u) \wedge\!\!\!\bigwedge_{q \in Q \setminus F} 
(q \rightarrow \bot)\]}
\begin{lemma}\label{lemma:abstract-invariant}
  For any bijection $\mu : \nat \rightarrow \Sigma^*$, there exists
  $k>0$ such that $\bigvee_{i=1}^k \AbsPost{\A}(\iota,\mu(i))$ defines
  an invariant for $\A$. 
\end{lemma}
\proof{ It is sufficient to show that there exists $k\geq0$ such that for
  all $u \in \Sigma^*$ there exists $i \in [0,k]$ such that
  $\Post{\A}(\iota,u) \models \AbsPost{\A}(\iota,\mu(i))$. We have
  $\Post{\A}(\iota,u) \models \AbsPost{\A}(\iota,u)$ for all $u \in
  \Sigma^*$. But since $\preds$ is a finite set, also the set
  $\{\AbsPost{\A}(\iota,u) \mid u \in \Sigma^*\}$ is finite. Thus
  there exists $k \geq 0$ such that, for all $u \in \Sigma^*$ there
  exists $i \in [0,k]$ such that $\AbsPost{\A}(\iota,u) \eqiff
  \AbsPost{\A}(\iota,\mu(i))$, which concludes the proof. \qed}

We are left with fulfilling the second point from the definition of a
safety invariant. To this end, suppose that, for a given set $\preds$
of predicates, the invariant $\abs{\Inv}$, defined by the previous
lemma, meets point (\ref{it1:inv}) but not point (\ref{it2:inv}) from
the definition of invariants, where $\Post{\A}$ and $\Accept{\A}$
replace $\AbsPost{\A}$ and $\AbsAccept{\A}$, respectively. In
other words, there exists a finite sequence $u \in \Sigma^*$ such that
$\nu \in \abs{\Inv}(\beta)$ and $\beta\cup\nu \models
\AbsAccept{\A}(u)$, for some boolean $\beta : Q \rightarrow \booli$
and data $\nu : \vec{x} \rightarrow \Data^\I$ valuations. Such
a $u\in\Sigma^*$ is called a \emph{counterexample}.

Once a counterexample $u$ is discovered, there are two
possibilities. Either\begin{inparaenum}[(i)]
\item $\Accept{\A}(u)$ is satisfiable, in which case $L(\A) \neq \emptyset$, or
\item $\Accept{\A}(u)$ is unsatisfiable, in which case $u$ is called
  \emph{spurious}.
\end{inparaenum}
In the first case, our semi-algorithm stops and returns a witness for
non-emptiness, obtained from the satisfying valuation of
$\Accept{\A}(u)$ and in the second case, we must strenghten the
invariant by excluding from $\abs{\Inv}$ all pairs $(\beta,\nu)$ such
that $\beta\cup\nu \models \AbsAccept{\A}(u)$. This strenghtening
is carried out by adding to $\preds$ several predicates that are
sufficient to exclude the spurious counterexample.

@@

\begin{lemma}\label{lemma:progress}
  Given a non-empty sequence of input events $u \in \Sigma^*$ such
  that $\Accept{\A}(u)$ is unsatisfiable, then the following
  hold: \begin{compactenum}
  \item\label{it1:progress} $\pi(u)$ is unsatisfiable, and
    %
  \item\label{it2:progress} if $\tuple{\top,I_1,\dots,I_n,\bot}$ is an
    interpolant for $\pi(u)$ and $I_i \in \preds$, for all
    $i\in[1,n]$, then $\AbsAccept{\A}(u)$ is unsatisfiable.
  \end{compactenum}
\end{lemma}
\proof{
